The Effects of the NBA COVID Bubble on the NBA Playoffs

### 2020-11-12 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

Data were collected from the official NBA website. The main variables of interest
are whether or not the home team won, scoring totals for home and away teams, and
shooting percentages for home and away teams on two pointers, three pointers, and
free throws. These variables were very popular and frequently used in the related
literature discussed earlier. While many other measures could be used for measuring
the outcome of the game and team performance, scoring seemed to be the most
important. The winner of a game is determined by who scores more points. So,
naturally when studying the outcome of winning, changes in scoring and scoring
efficiency are important to understand.

@@

A test for differences in means was conducted to compare scoring
totals. The null hypothesis is
\[
H_0: \mu_1 = \mu_2,
\]
where $\mu_1$ and $\mu_2$ are the mean scoring total in 2020 and
2017-2019, respectively. Let
$\bar{X}_1$ represent the average scoring total for 2020 and $\bar{X}_2$ represent
average scoring total over 2017-19. Similarly, let $s_1^2$ represent the sample variance
of scoring for 2020 and $s_2^2$ represent sample variance of scoring over 2017-19.
Finally, let $n_1$ represent the number of games in 2020 and $n_2$ represent the number
of games from 2017-2019. The null distribution of the test statistic is standard normal.
The test will be conducted twice once comparing home scoring in 2020 to 2017-19 and once
comparing away scoring in 2020 to 2017-2019. These tests will also be simaltaneously 
conducted using the Wilcoxon Test. The advantage of running the Wilcoxon Test is as a
nonparametric test it does not rely on any normality assumption

Tests for differences in proportions were conducted for comparing winning percentage
and the three different shooting percentages in 2020 to those of
2017-2019. The null hypothesis is 
\[
H_0: p_1 = p_2,
\]
where $p_1$ and $p_2$ are the shooting percentage in 2020 and
2017-2019, respectively. Let
$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
proportion for 2017-2019. The test statistic is
\begin{align}
    z = \frac{\hat{p}_1 - \hat{p}_2}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}},
    (\#eq:zprop)
\end{align}
where $\hat p$ is the pooled shooting percentage of 2017-2020.
The null distribution of the test statistic is also standard normal
The test will be conducted seven times,once comparing home winning percentage for 
2020 to that of 2017-2019 and once for each oftwo-point shooting, three-point shooting and
free throws for both home and away teams in 2020 vs. 2017-2019. Again these tests will
run simaltaneuously as a z-test and a Wilcoxon test.

>>>>

Data were collected from the official NBA website. The main variables of interest
are whether or not the home team won, scoring totals for home and away teams, and
shooting percentages for home and away teams on two pointers, three pointers, and
free throws. These variables were very popular and frequently used in the related
literature discussed earlier. The data was collected on a game by game basis, giving us one observation
for each variable per game played. There were 83 games played in the 2020 playoffs. Giving us 83 observations 
for each variable in 2020 for both home and away teams. Likewise, there was 243 games played over 
2017-2019. Giving us 243 observations of each variable for both home and away teams over 2017-2020. 
While many other measures could be used for measuring
the outcome of the game and team performance, scoring seemed to be the most
important. The winner of a game is determined by who scores more points. So,
naturally when studying the outcome of winning, changes in scoring and scoring
efficiency are important to understand.

@@

A test for differences in means was conducted to compare scoring
totals. The null hypothesis is
\[
H_0: \mu_1 = \mu_2,
\]
where $\mu_1$ and $\mu_2$ are the mean scoring total in 2020 and
2017-2019, respectively. Let
$\bar{X}_1$ represent the average scoring total for 2020 and $\bar{X}_2$ represent
average scoring total over 2017-19. Similarly, let $s_1^2$ represent the sample variance
of scoring for 2020 and $s_2^2$ represent sample variance of scoring over 2017-19.
Finally, let $n_1$ represent the number of games in 2020 and $n_2$ represent the number
of games from 2017-2019. The null distribution of the test statistic is standard normal.
The test will be conducted twice once comparing home scoring in 2020 to 2017-19 and once
comparing away scoring in 2020 to 2017-2019. These tests will also be simultaneously
conducted using the Wilcoxon Test. The advantage of running the Wilcoxon Test is as a
nonparametric test it does not rely on any normality assumption

Tests for differences in proportions were conducted for comparing winning percentage
and the three different shooting percentages in 2020 to those of
2017-2019. The null hypothesis is 
\[
H_0: p_1 = p_2,
\]
where $p_1$ and $p_2$ are the shooting percentage in 2020 and
2017-2019, respectively. Let
$\hat{p}_1$ represents sample proportion for 2020 and $\hat{p}_2$ represents sample
proportion for 2017-2019. The test statistic is
\begin{align}
    z = \frac{\hat{p}_1 - \hat{p}_2}{\sqrt{\hat{p}(1-\hat{p})(\frac{1}{n_1}+\frac{1}{n_2})}},
    (\#eq:zprop)
\end{align}
where $\hat p$ is the pooled shooting percentage of 2017-2020.
The null distribution of the test statistic is also standard normal
The test will be conducted seven times,once comparing home winning percentage for 
2020 to that of 2017-2019 and once for each oftwo-point shooting, three-point shooting and
free throws for both home and away teams in 2020 vs. 2017-2019. Again these tests will
run simultaneously as a z-test and a Wilcoxon test.
